\chapter*{Before You Begin}

\section*{\secfont What this manual covers}

This user manual provides details of how to download, install and
use \topspin, an automatic symmetry reduction tool for the \spin\
model checker.  \topspin\ can potentially aid in the verification of
\emph{safety} properties of concurrent systems specified in Promela.

\begin{itemize}
\item {\bf \downloadingandinstalling\ } \chapref{downloadandinstall}
provides details of other packages on which \topspin\ depends,
including where these packages can be found, and explains how to
download and install \topspin.

\item {\bf \workedexample\ } \chapref{example} provides a worked example showing how to run
\topspin\ to obtain symmetry reduction for an example specification.

\item {\bf \overviewofoptions\ } An overview of
\topspin\ options is presented in \chapref{overview}.

\item {\bf \buildingfromsource\ } For users who wish to experiment further with \topspin,
\chapref{compilingfromsource} explains how to obtain, compile and
test the \topspin\ source code.

\item {\bf \troubleshooting\ } \chapref{troubleshooting} presents solutions to common problems
associated with the installation and operation of \topspin, and
provides details of how bugs should be reported.

\end{itemize}

\section*{What this manual does not cover}

\begin{itemize}

\item {\bf Theory } Users interested in the theory on which
\topspin\ is based should refer to relevant papers and a Ph.D.
thesis, which are available from the tool web page (see below).

\item {\bf Use of \spin\ } This manual assumes that the reader is
familiar with the \spin\ tool and the Promela language.  Details of
and documentation for the \spin\ tool are available from the \spin\
web page.\footnote{\texttt{\spinurl}}

\end{itemize}

\section*{Online resources}

\begin{itemize}

\item {\bf \topspin\ web page } \texttt{\topspinwebpage}

\item {\bf SourceForge } \texttt{\sourceforgepage}

\end{itemize}
